\documentclass[openany,oneside]{amsbook}

\usepackage[all]{xy}
\usepackage[T1]{fontenc}
\usepackage{xstring}
\usepackage{xparse}
\usepackage{CJKutf8}
\usepackage[linktocpage=true,colorlinks=true,hyperindex,citecolor=blue,linkcolor=magenta]{hyperref}
\usepackage[margin=0.75in]{geometry}
\usepackage[charter,ttscaled=false,greekfamily=didot,uppercase=upright,greeklowercase=upright]{mathdesign}
\usepackage{tikz-cd}
\usepackage{enumitem}


\newtheoremstyle{ega-env-style}%
  {}{}{\rmfamily}{}{\bfseries}{.}{ }{\thmnote{(#3)}}%

\newtheoremstyle{ega-thm-env-style}%
  {}{}{\itshape}{}{\bfseries}{. --- }{ }{\thmname{#1}\thmnote{ (#3)}}%

\newtheoremstyle{ega-defn-env-style}%
  {}{}{\rmfamily}{}{\bfseries}{. --- }{ }{\thmname{#1}\thmnote{ (#3)}}%

\theoremstyle{ega-env-style}
\newtheorem*{env}{---}

\theoremstyle{ega-thm-env-style}
\newtheorem*{thm}{Theorem}
\newtheorem*{prop}{Proposition}
\newtheorem*{lem}{Lemma}
\newtheorem*{cor}{Corollary}

\theoremstyle{ega-defn-env-style}
\newtheorem*{defn}{Definition}
\newtheorem*{exm}{Example}
\newtheorem*{rmk}{Remark}
\newtheorem*{nota}{Notational conventions}

\makeatletter
\def\l@subsection{\@tocline{2}{0pt}{2.5pc}{5pc}{}}
\makeatother



\def\mathcal{\mathscr}
\def\sh{\mathcal}                 % sheaf font
\def\bb{\mathbf}                  % bold font
\def\cat{\mathtt}                 % category font
\def\fk{\mathfrak}                % mathfrak font
\def\wt{\widetilde}               % tilde over
\def\wh{\widehat}                 % hat over
\def\rad{\fk{r}}                  % radical
\def\nilrad{\fk{R}}               % nilradical
\def\emp{\varnothing}             % empty set
\def\vphi{\phi}                   % for switching \phi and \varphi, change if needed
\def\HH{\mathrm{H}}               % cohomology H
\def\CHH{\check{\HH}}             % Čech cohomology H
\def\dual#1{{#1}^\vee}            % dual
\def\kres{k}                      % residue field k
\def\C{\cat{C}}                   % category C
\def\op{^\cat{op}}                % opposite category
\def\Set{\cat{Set}}               % category of sets
\def\CHom{\cat{Hom}}              % functor category
\def\OO{\sh{O}}                   % structure sheaf O

\def\shHom{\sh{H}\!{om}}          % sheaf Hom
\def\shProj{\sh{P}\!{roj}}        % sheaf Proj
\def\rg{{\mathop{\mathrm{rg}}}}
\def\Hom{{\mathop{\mathrm{Hom}}\nolimits}}
\def\Proj{{\mathop{\mathrm{Proj}}\nolimits}}
\def\Tor{{\mathop{\mathrm{Tor}}\nolimits}}
\def\Ext{{\mathop{\mathrm{Ext}}\nolimits}}
\def\Supp{{\mathop{\mathrm{Supp}}\nolimits}}
\def\Ker{{\mathop{\mathrm{Ker}}\nolimits}\,}
\def\Im{{\mathop{\mathrm{Im}}\nolimits}\,}
\def\Coker{{\mathop{\mathrm{Coker}}\nolimits}\,}
\def\Spec{{\mathop{\mathrm{Spec}}\nolimits}}
\def\Spf{{\mathop{\mathrm{Spf}}\nolimits}}
\def\grad{{\mathop{\mathrm{grad}}\nolimits}}
\def\inf{{\mathop{\mathrm{inf}}\nolimits}}

\renewcommand{\to}{\mathchoice{\longrightarrow}{\rightarrow}{\rightarrow}{\rightarrow}}
\let\mapstoo\mapsto
\renewcommand{\mapsto}{\mathchoice{\longmapsto}{\mapstoo}{\mapstoo}{\mapstoo}}
\def\isoto{\simeq}  % isomorphism

\def\unsure#1{#1 {\color{red}(?)}}
\def\completelyunsure{{\color{red}(???)}}

\newcommand{\oldpage}[2][--]{{\marginpar{\textbf{#1}~|~#2}}\ignorespaces}
\def\sectionbreak{\begin{center}***\end{center}}

\NewDocumentCommand{\sref}{o m o}{%
  \IfNoValueTF{#1}%
    {\IfNoValueTF{#3}%
      {\hyperref[#2]{\normalfont{(\StrGobbleLeft{#2}{2})}}}%
      {\hyperref[#2]{\normalfont{(\StrGobbleLeft{#2}{2},~{#3})}}}}%
    {\IfNoValueTF{#3}%
      {\hyperref[#2]{\normalfont{(\textbf{#1},~\StrGobbleLeft{#2}{2})}}}%
      {\hyperref[#2]{\normalfont{(\textbf{#1},~\StrGobbleLeft{#2}{2},~{#3})}}}}%
}

\newcommand{\ZeroRoman}[1]{\ifcase\value{#1}\relax 0\else\Roman{#1}\fi}
\renewcommand{\thechapter}{\ZeroRoman{chapter}}
\begin{document}
\begin{titlepage}
\pagestyle{empty}
\setcounter{page}{1}
\centerline{\LARGE\bfseries \'El\'ements de g\'eom\'etrie alg\'ebrique}
\vskip1in
\noindent
\centerline{A.~Grothendieck and J.~Dieudonn\'e}
\centerline{Publications math\'ematiques de l'I.H.\'E.S}
\vskip1in
\noindent
\centerline{\bfseries Contributors}
\centerline{
Tim Hosgood, Ryan Keleti
}
\end{titlepage}
\setcounter{tocdepth}{2}
\tableofcontents

\chapter*{What this is}

\phantomsection
\label{section-phantom}

\noindent
\emph{This section is written by the translators.}

\noindent
This is a community translation of Grothendieck's \'El\'ements de g\'eom\'etrie alg\'ebrique (EGA).
As it is a work in progress by multiple people, there will probably be a few mistakes---if you spot any then please do \href{https://github.com/ryankeleti/en.ega/issues}{let us know}\footnote{\url{https://github.com/ryankeleti/en.ega/issues}}.

\noindent
To contribute, please visit
\begin{center}
  \url{https://github.com/ryankeleti/en.ega}.
\end{center}

\noindent
\emph{On est d\'esol\'es, Grothendieck.}

\section*{Notes from the translators}
Grothendieck's writing style in EGA is quite particular, most notably for its long sentence structure.
As translators, we have tried to give the best possible approximation of this style in English, resisting the temptation to ``streamline'' things in places where the language is more dense than usual.

\sectionbreak

Whenever a note is made by the translators, it will be prefaced by ``[Trans]''.

\sectionbreak

Along the margins we have provided the page numbers corresponding to the original text.
Due to EGA being a collection of volumes (one non-preliminary chapter per volume), the page numbers reset at every new chapter.
In addition, the preliminary section is stretched out over multiple volumes.
To combat this, we label the pages as
\begin{center}
  \textbf{X}~|~$n$,
\end{center}
referring to Chapter~X, page $n$.
In the case of the preliminaries, the preliminaries from volume~Y are denoted as \textbf{0\textsubscript{Y}}.

\sectionbreak

Later volumes (such as EGAs III and IV) include errata for earlier chapters.
Where possible, we have used these to `update' our translation, and entirely replace whatever mistakes might have been in the original copies of EGAs I and II, though we will try to include footnotes pointing out when this has taken place.

\section*{Mathematical notes}
EGA uses \emph{prescheme} for what is now usually called a scheme, and \emph{scheme} for what is now usually called a separated scheme.

\nocite{*}



\chapter*{Introduction}

\phantomsection
\label{section-phantom}


\begin{flushright}
\emph{To Oscar Zariski and Andr\'e Weil.}
\end{flushright}
\medskip

\oldpage[I]{5}
This memoir, and the many others will undoubtedly follow, are intended to form a treatise on the foundations of algebraic geometry.
They do not, in principle, presume any particular knowledge of the subject, and it has even been recognised that such knowledge, despite its obvious advantages, could sometimes (because of the much-too-narrow interpretation---through the birational point of view---that it usually implies) be a hindrance to the one who wants to become familiar with the point of view and techniques presented here.
However, we assume that the reader has a good knowledge of the following topics:
\begin{enumerate}[label=(\alph*)]
    \item \emph{Commutative algebra}, as it is laid out, for example, in the volumes (in progress of being written) of the \emph{\'El\'ements} of N.~Bourbaki (and, pending the publication of these volumes, in Samuel--Zariski \cite{I-13} and Samuel \cite{I-11,I-12}).
    \item \emph{Homological algebra}, for which we refer to Cartan--Eilenberg \cite{I-2} (cited as (M)) and Godement \cite{I-4} (cited as (G)), as well as the recent article by A. Grothendieck \cite{I-6} (cited as (T)).
    \item \emph{Sheaf theory}, where our main references will be (G) and (T);
        this theory provides the essential language for interpreting, in ``geometric'' terms, the essential notions of commutative algebra, and for ``globalizing'' them.
    \item Finally, it will be useful for the reader to have some familiarity with \emph{functorial language}, which will be constantly used in this treatise, and for which the reader may consult (M), (G), and especially (T);
        the principles of this language and the main results of the general theory of functors will be described in more detail in a book currently in preparation by the authors of this treatise.
\end{enumerate}

\sectionbreak

It is not the place, in this introduction, to give a more or less summarily description from the ``schemes'' point of view in algebraic geometry, nor the long list of reasons which made its adoption necessary, and in particular the systematic acceptance of nilpotent elements in the local rings of ``manifolds'' that we consider (which necessarily shifts the idea of rational maps into the background, in favor of those of regular maps or ``morphisms'').
To be precise, this treatise aims to systematically develop the language of schemes, and will demonstrate, we hope, its necessity.
Although it would be easy to do so,\oldpage[I]{6} we will not try to give here an ``intuitive'' introduction to the notions developed in Chapter~I.
For the reader who would like to have a glimpse of the preliminary study of the subject matter of this treatise, we refer them to the conference by A.~Grothendieck at the International Congress of Mathematicians in Edinburgh in 1958 \cite{I-7}, and the expos\'e \cite{I-8} of the same author.
The work \cite{I-14} (cited as (FAC)) of J.-P.~Serre can also be considered as an intermediary exposition between the classical point of view and the schemes point of view in algebraic geometry, and, as such, its reading may be an excellent preparation for the reading of our \emph{\'El\'ements}.

\sectionbreak

We give below the general outline planned for this treatise, subject to later modifications, especially concerning the later chapters.

\begin{tabular}{rrl}
Chapter & I. & --- The language of schemes.\\
--- & II. & --- Elementary global study of some classes of morphisms.\\
--- & III. & --- Cohomology of algebraic coherent sheaves. Applications.\\
--- & IV. & --- Local study of morphisms.\\
--- & V. & --- Elementary procedures of constructing schemes.\\
--- & VI. & --- Descent. General method of constructing schemes.\\
--- & VII. & --- Schemes of groups, principal fibre bundles.\\
--- & VIII. & --- Differential study of fibre bundles.\\
--- & IX. & --- The fundamental group.\\
--- & X. & --- Residues and duality.\\
--- & XI. & --- Theories of intersection, Chern classes,
Riemann--Roch theorem.\\
--- & XII. & --- Abelian schemes and Picard schemes.\\
--- & XIII. & --- Weil cohomology.
\end{tabular}\\

\bigskip

In principle, all chapters are considered open to changes, and supplementary sections could always be added later;
such sections would appear in separate fascicles in order to minimize the inconveniences accompanying whatever mode of publication adopted.
When the writing of such a section is foreseen or in progress during the publication of a chapter, it will be mentioned in the summary of the chapter in question, even if, owing to certain orders of urgency, its actual publication clearly ought to have been later.
For the convenience of the reader, we give in ``Chapter~0'' the necessary tools in commutative algebra, homological algebra, and sheaf theory, that will be used throughout this treatise, that are more or less well known but for which it was not possible to give convenient references.
It is recommended for the reader to not read Chapter~0 except whilst reading the actual treatise, when the results to which we refer\oldpage[I]{7} seem unfamiliar.
Besides, we think that in this way, the reading of this treatise could be a good method for the beginner to familiarize themselves with commutative algebra and homological algebra, whose study, when not accompanied with tangible applications, is considered tedious, or even depressing, by many.

\sectionbreak

It is outside of our capabilities to give a historic overview, or even a summary thereof, of the ideas and results described herein.
The text will contain only those references considered particularly useful for comprehension, and we indicate the origin of only the most important results.
Formally, at least, the subjects discussed in our work are reasonably new, which explains the scarcity of references made to the fathers of algebraic geometry from the 19th to the beginning of the 20th century, whose works we know only by hear-say.
It is suitable, however, to say some words here about the works which have most directly influenced the authors and contributed to the development of scheme-theoretic point of view.
We absolutely must mention the fundamental work (FAC) of J.-P.~Serre first, which has served as an introduction to algebraic geometry for more that one young student (the author of this treatise being one), deterred by the dryness of the classic \emph{Foundations} of A.~Weil \cite{I-18}.
It is there that it is shown, for the first time, that the ``Zariski topology'' of an ``abstract'' algebraic variety is perfectly suited to applying certain techniques from algebraic topology, and notably to be able to define a cohomology theory.
Further, the definition of an algebraic variety given therein is that which translates most naturally to the idea that we develop here\footnote{Just as J.-P.~Serre informed us, it is right to note that the idea of defining the structure of a manifold by the data of a sheaf of rings is due to H.~Cartan, who took this idea as the starting point of his theory of analytic spaces.%
Of course, just as in algebraic geometry, it would be important in ``analytic geometry'' to give the allow the use of nilpotent elements in local rings of analytic spaces.%
This extension of the definition of H.~Cartan and J.-P.~Serre has recently been broached by H.~Grauert \cite{I-5}, and there is room to hope that a systematic report of analytic geometry in this setting will soon see the light of day.%
It is also evident that the ideas and techniques developed in this treatise retain a sense of analytic geometry, even though one must expect more considerable technical difficulties in this latter theory.%
We can foresee that algebraic geometry, by the simplicity of its methods, will be able to serve as a sort of formal model for future developments in the theory of analytic spaces.}.
Serre himself had incidentally noted that the cohomology theory of affine algebraic varieties could be translated without difficulty by replacing the affine algebras over a field by arbitrary commutative rings.
Chapters~I and II of this treatise, and the first two paragraphs of Chapter~III, can thus be considered, for the most part, as easy translations, to this bigger framework, of the principal results of (FAC) and a later article of the same author \cite{I-15}.
We have also vastly profited from the \emph{S\'eminaire de G\'eom\'etrie Alg\'ebrique} de C.~Chevalley \cite{I-1};
in particular, the systematic usage of ``constructible sets'' introduced by him has turned out to be extremely useful in the theory of schemes (cf. Chapter~IV).
We have also borrowed from him the study of morphisms from\oldpage[I]{8} the point of view of dimension (Chapter~IV), that translates with negligible change to the framework of schemes.
It also merits noting that the idea of ``schemes of local rings'', introduced by Chevalley, naturally lends itself to being extended to algebraic geometry (not having, however, all the flexibility and generality that we intend to give it here);
for the connections between this idea and our theory, see Chapter~I, \textsection8.
One such extension has been developed by M.~Nagata in a series of memoirs \cite{I-9}, which contain many special results concerning algebraic geometry over Dedekind rings\footnote{Among the works that come close to our point of view of algebraic geometry, we pick out the work of E.~K\"ahler \cite{I-22} and a recent note of Chow and Igusa \cite{I-3}, which go back over certain results of (FAC) in the context of Nagata--Chevalley theory, as well as giving a K\"unneth formula.}.

\sectionbreak

It goes without saying that a book on algebraic geometry, and especially a book dealing with the fundamentals, is of course influenced, if only by proxy, by mathematicians such as O.~Zariski and A.~Weil.
In particular, the \emph{Th\'eorie des fonctions holomorphes} by Zariski \cite{I-20}, reasonably flexible thanks to the cohomological methods and an existence theorem (Chapter~III, \textsection\textsection4 and 5), is (along with the method of descent described in Chapter~VI) one of the principal tools used in this treatise, and it seems to us one of the most powerful at our disposal in algebraic geometry.

The general technique in which it is employed can be sketched as follows (a typical example of which will be given in Chapter~XI, in the study of the fundamental group).
We have a proper morphism (Chapter~II) $f:X\to Y$ of algebraic varieties (or, more generally, of schemes) that we wish to study on the neighborhood of a point $y\in Y$, with the aim of resolving a problem $P$ relative to a neighborhood of $y$.
We proceed step by step:
\begin{enumerate}
  \item[1st]
    We can suppose that $Y$ is affine, so that $X$ becomes a scheme defined on the affine ring $A$ of $Y$, and we can even replace $A$ by the local ring of $y$.
    This reduction is always easy in practice (Chapter~V) and brings us to the case where $A$ is a \emph{local} ring.
  \item[2nd]
    We study the problem in question when $A$ is a local \emph{Artinian} ring.
    So that the problem $P$ still makes sense when $A$ is not assumed to be integral, we sometimes have to reformulate $P$, and it appears that we often obtain a better understanding of the problem in doing so, in an ``infinitesimal'' way.
  \item[3rd]
    The theory of formal schemes (Chapter~III, \textsection\textsection3, 4, and 5) lets us pass from the case of an Artinian ring to a \emph{complete local ring}.
  \item[4th]
    Finally, if $A$ is an arbitrary local ring, considering ``\unsure{multiform} sections'' of suitable schemes over $X$, approximating a given ``formal'' section (Chapter~IV), will let us pass\oldpage[I]{9}, by extension of scalars, to the completion of $A$, from a known result (about the schemed induced by $X$ by extension of scalars to the completion of $A$) to an analogous result for a finite simple (e.g. unramified) extension of $A$.
\end{enumerate}

This sketch shows the importance of the systematic study of schemes defined over an Artinian ring $A$.
The point of view of Serre in his formulation of the theory of local class fields, and the recent works of Greenberg, seem to suggest that such a study could be undertaken by functorially attaching, to some such scheme $X$, a scheme $X'$ over the residue field $k$ of $A$ (assumed perfect) of dimension equal (in nice cases) to $n\dim X$, where $n$ is the height of $A$.

As for the influence of A.~Weil, it suffices to say that it is the need to develop the tools necessary to formulate, with full generality, the definition of ``Weil cohomology'', and to tackle the proof\footnote{To avoid any misunderstanding, we point out that this task has barely been undertaken at the moment of writing this introduction, and still hasn't led to the proof of the Weil conjectures.} of all the formal properties necessary to establish the famous conjectures in Diophantine geometry \cite{I-19}, that has been one of the principal motivations for the writing of this treatise, as well as the desire to find the natural setting of the usual ideas and methods of algebraic geometry, and to give the authors the chance to understand said ideas and methods.

\sectionbreak

Finally, we believe it useful to warn the reader that they, as did all the authors themselves, will almost certainly have difficulty before becoming accustomed to the language of schemes, and to convince themselves that the usual constructions that suggest geometric intuition can be translated, in essentially only one sensible way, to this language.
As in many parts of modern mathematics, the first intuition seems further and further away, in appearance, from the correct language needed to express the mathematics in question with complete precision and the desired level of generality.
In practice, the psychological difficulty comes from the need to replicate some familiar set-theoretic constructions to a category that is already quite different from that of sets (the category of preschemes, or the category of preschemes over a given prescheme): Cartesian products, group laws, ring laws, module laws, fibre bundles, principal homogeneous fibre bundles, etc.
It will most likely be difficult for the mathematician, in the future, to shy away from this new effort of abstraction (maybe rather negligible, on the whole, in comparison with that supplied by our fathers) to familiarize themselves with the theory of sets.

\sectionbreak

The references are given following the numerical system; for example, in III,~4.9.3, the III indicates the volume, the 4 the chapter, the 9 the section, and the 3 the paragraph.\footnote{\emph{[Trans] This is not a direct translation of the original, but instead uses the language more familiar to modern book (and LaTeX document) layouts.}}
If we reference a volume from within itself then we omit the volume number.

\bigskip

\oldpage[I]{10}
\emph{[Trans] Page 10 in the original is left blank.}




\setcounter{chapter}{-1}\chapter{Preliminaries}

\phantomsection
\label{section-phantom}


\input{ega0/ega0-1}
\input{ega0/ega0-2}
\input{ega0/ega0-3}
\input{ega0/ega0-4}
\input{ega0/ega0-5}
\input{ega0/ega0-6}
\input{ega0/ega0-7}
\input{ega0/ega0-8}




\chapter{The language of schemes (EGA~I)}

\phantomsection
\label{section-phantom}


\section*{Summary}
\label{section-egaI-summary}

\begin{tabular}{ll}
  \textsection1. & Affine schemes.\\
  \textsection2. & Preschemes and morphisms of preschemes.\\
  \textsection3. & Products of preschemes.\\
  \textsection4. & Subpreschemes and immersion morphisms.\\
  \textsection5. & Reduced preschemes; separation condition.\\
  \textsection6. & Finiteness conditions.\\
  \textsection7. & Rational maps.\\
  \textsection8. & Chevalley schemes.\\
  \textsection9. & Supplement on quasi-coherent sheaves.\\
  \textsection10. & Formal schemes.
\end{tabular}\\

\oldpage[I]{79}
The \textsection\textsection1--8 do little more than develop a language, which will be used in the following.
It should be noted, however, that in accordance with the general spirit of this treatise, \textsection\textsection7--8 will be used less than the others, and in a less essential way; we have moreover spoken of Chevalley's schemes only to make the link with the language of Chevalley \cite{I-1} and Nagata \cite{I-9}.
The \textsection9 gives definitions and results on quasi-coherent sheaves, some of which are no longer limited to a translation into a ``geometric'' language of known notions of commutative algebra, but are already of a global nature; they will be indispensable, in the following chapters, for the global study of morphisms.
Finally, \textsection10 introduces a generalization of the notion of schemes, which will be used as an intermediary in Chapter~III to formulate and prove in a convenient way the fundamental results of the cohomological study of the proper morphisms; moreover, it should be noted that the notion of formal schemes seems indispensable to express certain facts of the ``theory of modules'' (classification problems of algebraic varieties).
The results of \textsection10 will not be used before \textsection3 of Chapter~III and it is recommended to omit reading until then.
\bigskip

\input{egaI/egaI-1}
\input{egaI/egaI-2}
\input{egaI/egaI-3}
\input{egaI/egaI-4}
\input{egaI/egaI-5}
\input{egaI/egaI-6}
\input{egaI/egaI-7}
\input{egaI/egaI-8}
\input{egaI/egaI-9}
\input{egaI/egaI-10}




\chapter{Elementary global study of some classes of morphisms (EGA~II)}

\phantomsection
\label{section-phantom}

build hack
\cite{I-1}


\section*{Summary}
\label{section-egaII-summary}

\begin{tabular}{ll}
  \textsection1. & Affine morphisms\\
  \textsection2. & Homogeneous prime spectra.\\
  \textsection3. & Homogeneous prime spectrum of a sheaf of graded algebras.\\
  \textsection4. & Projective bundles; ample sheaves.\\
  \textsection5. & Quasi-affine morphisms; quasi-projective morphisms; proper morphisms; projective morphisms.\\
  \textsection6. & Integral morphisms and finite morphisms.\\
  \textsection7. & Valuative criteria.\\
  \textsection8. & Blowup schemes; projective cones; projective closure.\\
\end{tabular}\\

\oldpage[II]{5}
The various classes of morphisms studied in this chapter are used extensively in cohomological methods; further study, using these methods, will be done in Chapter~III, where we use especially \textsection\textsection2,4, and 5 of Chapter~II.
The \textsection8 can be omitted on a first reading: it gives some supplements to the formalism developed in \textsection\textsection1 and 3, reducing to easy maps of this formalism, and we will use it less consistently than the other results of this chapter.
\bigskip

\input{egaII/egaII-1}
\input{egaII/egaII-2}
\input{egaII/egaII-3}
\input{egaII/egaII-4}
\input{egaII/egaII-5}
\input{egaII/egaII-6}
\input{egaII/egaII-7}
\input{egaII/egaII-8}




\chapter{Cohomological study of coherent sheaves (EGA~III)}

\phantomsection
\label{section-phantom}

build hack
\cite{I-1}


\section*{Summary}
\label{section-egaIII-summary}

\begin{tabular}{ll}
    \textsection1. & Cohomology of affine schemes.\\
    \textsection2. & Cohomological study of projective morphisms.\\
    \textsection3. & Finiteness theorem for proper morphisms.\\
    \textsection4. & The fundamental theorem of proper morphisms. Applications.\\
    \textsection5. & An existence theorem for coherent algebraic sheaves.\\
    \textsection6. & Local and global Tor functors; Künneth formula.\\
    \textsection7. & Base change for homological functors of sheaves of modules.\\

    \textsection8. & The duality theorem for projective bundles\\
    \textsection9. & Relative cohomology and local cohomology; local duality\\
    \textsection10. & Relations between projective cohomology and local cohomology. Formal completion technique along a divisor\\
    \textsection11. & Global and local Picard groups\footnote{EGA IV does not depend on \textsection\textsection8 to 11, and will probably be published before these chapters. \emph{[Trans.] these last four chapters were never published.}}
\end{tabular}\\

This chapter gives the fundamental theorems concerning the cohomology of coherent algebraic sheaves, with the exception of theorems explaining the theory of residues (duality theorems), which will be the subject of a later chapter.
Amongst those included, there are essentially six fundamental theorems, each one being the subject of one of the first six chapters.
These results will prove to be essential tools in all that follows, even in questions which are not truly cohomological in their nature;
the reader will see the first such examples starting from \textsection4.
Section \textsection7 gives some more technical results, but which are used constantly in applications.
Finally, in \textsection\textsection8 to 11, we will develop certain results, linked to the duality of coherent sheaves, that are particularly important for applications, and which can show up before the general theory of residues.

The contents of \textsection\textsection1 and 2 is due to J.-P.~Serre, and the reader will observe that we have had only to follow (FAC).
Sections \textsection8 and 9 are equally inspired by (FAC) (the changes necessitated by the different contexts, however, being less evident).
Finally, as we said in the Introduction, section \textsection4 should be considered as the formatting, in modern language, of the fundamental ``invariance theorem'' of Zariski's ``theory of holomorphic functions''.

We draw attention to the fact that the results of n\textsuperscript{o}~3.4 (and the preliminary propositions of (\textbf{0},~13.4 to 13.7)) will not be used in what follows chap.~III, and can thus be skipped in a first reading.

\bigskip

\input{egaIII/egaIII-1}
\input{egaIII/egaIII-2}
\input{egaIII/egaIII-3}
\input{egaIII/egaIII-4}
\input{egaIII/egaIII-5}
\input{egaIII/egaIII-6}
\input{egaIII/egaIII-7}



\bibliography{the}
\bibliographystyle{amsalpha}
\end{document}
