\documentclass[../main.tex]{subfiles}

\begin{document}

\begin{cx}{1.5.1}
Let $A$, $A'$ be two rings, $\varphi$ a homomorphism $A'\to A$, $S$ (resp. $S'$)
a multiplicative subset of $A$ (resp. $A'$), such that $\varphi(S')\su S$; the
composition homomorphism $\smash{A'\xrightarrow{\varphi} A\to S^{-1}A}$ factors as
$\smash{A'\to{S'}^{-1}\xrightarrow{\varphi^{S'}} S^{-1}A}$ by virtue of (1.2.4);
where $\varphi^{S'}(a'/s')=\varphi(a')/\varphi(s')$. If $A=\varphi(A')$ and
$S=\varphi(S')$, $\varphi^{S'}$ is \emph{surjective}. If $A'=A$ and if $\varphi$
is the identity, $\varphi^{S'}$ is none other than the homomorphism $\rho_A^{S,S'}$
defined in (1.4.1).
\end{cx}

\begin{cx}{1.5.2}
Under the hypothesis of (1.5.1), let $M$ be an $A$-module. There exists a canonical
functorial morphism
\[\begin{tikzcd}
  \sigma:{S'}^{-1}(M_{[\varphi]})\ar[r] & (S^{-1}M)_{[\varphi^{S'}]}
\end{tikzcd}\]
of ${S'}^{-1}A'$-modules, sending each element $m/s'$ of ${S'}^{-1}(M_{[\varphi]})$ to
the element $m/\varphi(s')$ of $(S^{-1}M)_{[\varphi^{S'}]}$; in fact, we verify
immediately that this definition does not depend on the expression $m/s'$ of the element
considered. \emph{When} $S=\varphi(S')$, the homomorphism $\sigma$ is \emph{bijective}.
When $A'=A$ and $\varphi$ is the identity, $\sigma$ is none other that the homomorphism
$\rho_M^{S,S'}$ defined in (1.4.1).

When $M=A$ is taken in particular, the homomorphism $\varphi$ defines on $A$ an $A'$-algebra
structure; ${S'}^{-1}(A_{[\varphi]})$ is then provided with a ring structure, for which it
identifies with $(\varphi(S'))^{-1}A$, and the homomorphism
${\sigma:{S'}^{-1}(A_{[\varphi]})\to S^{-1}A}$ is a homomorphism of ${S'}^{-1}A'$-algebras.
\end{cx}

\begin{cx}{1.5.3}
Let $M$ and $N$ be two $A$-modules; by composing the homomorphisms defined in (1.3.4) and
(1.5.2), we obtain a homomorphism
\[\begin{tikzcd}
  (S^{-1}M\otimes_{S^{-1}A}S^{-1}N)_{[\varphi^{S'}]} & {S'}^{-1}((M\otimes A)_{[\varphi]})\ar[l]
\end{tikzcd}\]
which is an isomorphism when $\varphi(S')=S$. Similarly, by composing the homorphisms in (1.3.5)
and (1.5.2), we obtain a homomorphism
\[\begin{tikzcd}
  {S'}^{-1}((\Hom_A(M,N))_{[\varphi]})\ar[r] & (\Hom_{S^{-1}A}(S^{-1}M,S^{-1}N))_{[\varphi^{S'}]}
\end{tikzcd}\]
which is an isomorphism when $\varphi(S')=S$ and $M$ admits a finite presentation.
\end{cx}

\end{document}

