\documentclass[../main.tex]{subfiles}

\begin{document}

\begin{cx}[1.0.1]
All the rings considered in this Treatise will have a \emph{unit element}; all the modules
on such a ring will be assumed to be \emph{unitary}; the ring homomorphisms will always be
assumed to \emph{transform the unit element into a unit element}; unless otherwise stated,
a sub-ring of a ring $A$ will be assumed to \emph{contain the unit element of} $A$. We will
consider especially \emph{commutative} rings, and when we speak of a ring without
specification, it
will be implied that it is commutative. If $A$ is a ring not necessarily commutative, by
$A$-module we will we mean a left module, unless stated otherwise.
\end{cx}

\begin{cx}[1.0.2]
Let $A$, $B$ be two rings, not necessarily commutative, $\varphi:A\to B$ a homomorphism.
Any left (resp. right) $B$-module $M$ can be provided with a left (resp. right) $A$-module
structure by $a\cdot m=\varphi(a)\cdot m$ (resp. $m\cdot a=m\cdot\varphi(a)$); when it will
be necessary to distinguish $M$ as an $A$-module or a $B$-module, we will denote by
$M_{[\varphi]}$ the left (resp. right) $A$-module as defined. If $L$ is an $A$-module, then
a homomorphism $u:L\to M_{[\varphi]}$ is a homomorphism of commutative groups such that
$u(a\cdot x)=\varphi(a)\cdot u(x)$ for $a\in A$, $x\in L$; we will also say that it is a
$\varphi$-\emph{homomorphism} $L\to M$,
and that the pair $(\varphi,u)$ (or, by misuse of langauge, $u$)
is a \emph{di-homomorphism} of $(A,L)$ in $(B,M)$. The pairs $(A,L)$ formed by a ring $A$
and an $A$-module $L$ thus form a \emph{category} for which the morphisms are
di-homomorphisms.
\end{cx}

\begin{cx}[1.0.3]
Under the hypothesis of (1.0.2), if $\mf{J}$ is a left (resp. right) ideal of $A$, we
denote by $B\mf{J}$ (resp. $\mf{J}B$) the left (resp. right) ideal $B\varphi(\mf{J})$
(resp. $\varphi(\mf{J})B$) of $B$ generated by $\varphi(\mf{J})$; it is also the image
of the canonical homomorphism $B\otimes_A\mf{J}\to B$ (resp. $\mf{J}\otimes_A B\to B$)
of left (resp. right) $B$-modules.
\end{cx}

\end{document}

